Analysis-ready VCF at Biobank scale using Zarr

### 2024-11-14 paper.tex

%% Regarding data storage files, which to use?
%% find spruce.icf -name "*" | wc -l -> 1394080
%% # chunks in spruce.icf.inspect.txt -> 1342569
%% du -hs spruce.icf -> 6.8TiB
%% sum size in spruce.icf.inspect.txt -> 6.8TiB
%% du -hs spruce.vcz -> 6.7TiB
%% sum size in spruce.vcz.inspect.txt -> 6.6TiB
%% # chunks in spruce.vcz.inspect.txt -> 11984581
%% More chunk files despite smaller input data: could selection of
%% chunk size matter here?
Following the conversion process detailed in the previous section, we
first converted the 165 VCF files (7.4TiB) to ICF, followed by
conversion to Zarr. The tasks were run as single jobs on a compute
node with 2,048 GB RAM and 128 cores. Conversion to ICF required
13h45min runtime. The ICF representation used a total of 6.8 TiB over
1,394,080 data storage files. Conversion to Zarr required 18h51min
runtime, generating a dataset with 32 arrays, consuming a total of 6.7
TiB storage over 11,984,581 chunk files. The reduction in total
storage (1.1X) is negligible, presumably due to the small sample size.
The top fields in terms of storage are detailed in
Table~\ref{tab-spruce-data}.

\begin{table}
\caption{Summary for a selection of the largest VCF Zarr columns produced for 
1,063 VCFs on 12 spruce chromosomes using \texttt{vcf2zarr}
default settings. For column details see the caption of Table~\ref{tab-genomics-england-data}.
\label{tab-spruce-data}}
\begin{tabular}{llS[table-format=3.1]S[table-format=3.2]S[table-format=3.2]}
\toprule
{Field} & {type} & {storage} & {compress} & {\%total} \\
  \midrule
  /call\_PL &	int16 &	6.1T &	12 & 92.75\%\\
  /call\_genotype & int8 & 282.5G & 26 & 4.18\%\\
  /call\_genotype\_mask & bool& 42.2G & 180 & 0.62\%\\
  /variant\_DP4 & int32 & 26.2G & 2.1 & 0.39\%\\
  /variant\_MQSBZ & float32 & 13.1G & 1.1 & 0.19\%\\
  /variant\_RPBZ & float32 & 13.0G & 1.1 & 0.19\%\\
  /variant\_MQBZ & float32 & 13.0G & 1.1 & 0.19\%\\
  /variant\_SCBZ & float32 & 12.9G & 1.1 & 0.19\%\\
  /variant\_VDB & float32 & 12.8G & 1.1 & 0.19\%\\
  /variant\_quality & float32 &12.8G & 1.1 & 0.19\%\\
  /variant\_BQBZ & float32 & 12.6G & 1.1 & 0.19\%\\
  /variant\_SGB & float32 & 12.1G & 1.2 & 0.18\%\\
  /variant\_position & int32 & 9.7G & 1.4 & 0.14\%\\
  /variant\_AC & int16 & 7.4G & 2.8 & 0.11\%\\
  /variant\_DP & int32 & 6.7G & 2.1 & 0.10\%\\
  /variant\_allele & object & 5.4G & 21 & 0.08\%\\
  /variant\_AN & int16 & 3.1G & 2.3 & 0.05\%\\
  /call\_genotype\_phased & bool & 2.0G & 1800 & 0.03\%\\
  /variant\_filter & bool & 1.5G & 2.4 & 0.02\%\\
  /variant\_MQ & int8 &	489.8M & 7.3 &	0.01\%\\
%   "/variant_IMF"	"float32"	"161.37 MiB"	89	"0.00%"
% "/variant_IDV"	"int16"	"121.14 MiB"	59	"0.00%"
% "/variant_INDEL"	"bool"	"72.6 MiB"	49	"0.00%"
% "/variant_id"	"object"	"34.95 MiB"	820	"0.00%"
% "/variant_id_mask"	"bool"	"33.16 MiB"	110	"0.00%"
% "/variant_contig"	"int16"	"32.44 MiB"	220	"0.00%"
% "/variant_FS"	"float32"	"32.09 MiB"	450	"0.00%"
% "/variant_MQ0F"	"float32"	"32.09 MiB"	450	"0.00%"
% "/contig_length"	"int64"	"6.77 KiB"	1.2	"0.00%"
% "/contig_id"	"object"	"6.2 KiB"	1.3	"0.00%"
% "/sample_id"	"object"	"5.46 KiB"	1.5	"0.00%"
% "/filter_id"	"object"	"4.43 KiB"	0	"0.00%"
  \bottomrule
\end{tabular}
\end{table}

>>>>

%% Regarding data storage files, which to use?
%% find spruce.icf -name "*" | wc -l -> 1394080
%% # chunks in spruce.icf.inspect.txt -> 1342569
%% du -hs spruce.icf -> 6.8TiB
%% sum size in spruce.icf.inspect.txt -> 6.8TiB
%% du -hs spruce.vcz -> 6.7TiB
%% sum size in spruce.vcz.inspect.txt -> 6.6TiB
%% # chunks in spruce.vcz.inspect.txt -> 11984581
%% More chunk files despite smaller input data: could selection of
%% chunk size matter here?
Following the conversion process detailed in the previous section, we
first converted the 165 VCF files (7.33TiB) to ICF, followed by
conversion to Zarr. The tasks were run as single jobs on a compute
node with 2,048 GB RAM and 128 cores. Conversion to ICF required
13h45min runtime. The ICF representation used a total of 6.77 TiB over
1,342,569 data storage files. Conversion to Zarr required 18h51min
runtime, generating a dataset with 32 arrays, consuming a total of 6.6
TiB storage over 11,984,581 chunk files. The reduction in total
storage (1.1X) is negligible, mainly due to the inherent difficulty in
compressing the PL field. The top fields in terms of storage are
detailed in Table~\ref{tab-spruce-data}.

\begin{table}
\caption{Summary for a selection of the largest VCF Zarr columns produced for 
1,063 VCFs on 12 spruce chromosomes using \texttt{vcf2zarr}
default settings. For column details see the caption of Table~\ref{tab-genomics-england-data}.
\label{tab-spruce-data}}
\begin{tabular}{llS[table-format=3.1]S[table-format=3.2]S[table-format=3.2]}
\toprule
{Field} & {type} & {storage} & {compress} & {\%total} \\
  \midrule
  /call\_PL & int16 & 6.12 TiB & 12.0 & 92.75\% \\
  /call\_genotype & int8 & 282.45 GiB & 26.0 & 4.18\% \\
  /call\_genotype\_mask & bool & 42.15 GiB & 180.0 & 0.62\% \\
  /variant\_DP4 & int32 & 26.24 GiB & 2.1 & 0.39\% \\
  /variant\_MQSBZ & float32 & 13.08 GiB & 1.1 & 0.19\% \\
  /variant\_RPBZ & float32 & 13.04 GiB & 1.1 & 0.19\% \\
  /variant\_MQBZ & float32 & 12.97 GiB & 1.1 & 0.19\% \\
  /variant\_SCBZ & float32 & 12.94 GiB & 1.1 & 0.19\% \\
  /variant\_VDB & float32 & 12.84 GiB & 1.1 & 0.19\% \\
  /variant\_quality & float32 & 12.84 GiB & 1.1 & 0.19\% \\
  /variant\_BQBZ & float32 & 12.64 GiB & 1.1 & 0.19\% \\
  /variant\_SGB & float32 & 12.12 GiB & 1.2 & 0.18\% \\
  /variant\_position & int32 & 9.65 GiB & 1.4 & 0.14\% \\
  /variant\_AC & int16 & 7.42 GiB & 2.8 & 0.11\% \\
  /variant\_DP & int32 & 6.71 GiB & 2.1 & 0.10\% \\
  /variant\_allele & object & 5.37 GiB & 21.0 & 0.08\% \\
  /variant\_AN & int16 & 3.05 GiB & 2.3 & 0.05\% \\
  /call\_genotype\_phased & bool & 2.03 GiB & 1800.0 & 0.03\% \\
  /variant\_filter & bool & 1.46 GiB & 2.4 & 0.02\% \\
  /variant\_MQ & int8 & 489.83 MiB & 7.3 & 0.01\% \\
  \bottomrule
\end{tabular}
\end{table}
